### 2022-10-31 docs/paper.tex
init docs

\documentclass[
oneside,
fontsize=11pt
]{scrartcl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% packages
%%%

%%%
%%% encoding and language set
%%%

\usepackage[english]{babel}
\usepackage[a4paper]{geometry}

%%% fontenc, ae, aecompl: coding of characters in PDF documents
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage[autostyle=true,german=quotes]{csquotes}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{minted}

%%%
%%% technical packages
%%%

%%% amsmath, amssymb, amstext: support for mathematics
\usepackage{amsmath,amssymb,amstext}

%%%
%%% layout
%%%

%%% scrpage2: KOMA heading and footer
%%% Note: if you don't use this package, please remove 
%%%       \pagestyle{scrheadings} and corresponding settings
%%%       below too.
\usepackage[automark]{scrpage2}

%%%
%%% Colors
%%%

\RequirePackage[table,dvipsnames]{xcolor}

\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}

%%%
%%% PDF
%%%

\usepackage{ifpdf}

%%% Should be LAST usepackage-call!
%%% For docu on that, see reference on package ``hyperref''
\ifpdf%   (definitions for using pdflatex instead of latex)

  %%% graphicx: support for graphics
  \usepackage[pdftex]{graphicx}

  \pdfcompresslevel=9

  %%% hyperref (hyperlinks in PDF): for more options or more detailed
  %%%          explanations, see the documentation of the hyperref-package
  \usepackage[%
    %%% general options
    pdftex=true,      %% sets up hyperref for use with the pdftex program
    %plainpages=false, %% set it to false, if pdflatex complains: ``destination with same identifier already exists''
    %
    %%% extension options
    backref,      %% adds a backlink text to the end of each item in the bibliography
    pagebackref=false, %% if true, creates backward references as a list of page numbers in the bibliography
    colorlinks=true,   %% turn on colored links (true is better for on-screen reading, false is better for printout versions)
    linkcolor=TUGreen,
    urlcolor=TUGreen,
    %
    %%% PDF-specific display options
    bookmarks=true,          %% if true, generate PDF bookmarks (requires two passes of pdflatex)
    bookmarksopen=true,     %% if true, show all PDF bookmarks expanded
    bookmarksnumbered=true, %% if true, add the section numbers to the bookmarks
    %pdfstartpage={1},        %% determines, on which page the PDF file is opened
    pdfpagemode=None         %% None, UseOutlines (=show bookmarks), UseThumbs (show thumbnails), FullScreen
  ]{hyperref}

  %%% provide all graphics (also) in this format, so you don't have
  %%% to add the file extensions to the \includegraphics-command
  %%% and/or you don't have to distinguish between generating
  %%% dvi/ps (through latex) and pdf (through pdflatex)
  \DeclareGraphicsExtensions{.pdf}

\else %else   (definitions for using latex instead of pdflatex)

  \usepackage[dvips]{graphicx}

  \DeclareGraphicsExtensions{.eps}

  \usepackage[%
    dvips,           %% sets up hyperref for use with the dvips driver
    colorlinks=false %% better for printout version; almost every hyperref-extension is eliminated by using dvips
  ]{hyperref}

\fi

%%% sets the PDF-Information options
%%% (see fields in Acrobat Reader: ``File -> Document properties -> Summary'')
%%% Note: this method is better than as options of the hyperref-package (options are expanded correctly)
\hypersetup{
  pdftitle={Fréchet Distance}, %%
  pdfauthor={Tom Stein}, %%
  pdfsubject={Seminar Algorithm Engineering 22/23}, %%
  pdfcreator={Accomplished with LaTeX2e and pdfLaTeX with hyperref-package.}, %% 
  pdfproducer={}, %%
  pdfkeywords={} %%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% define the titlepage
%%%

% \subject{}   %% subject which appears above titlehead
% \titlehead{} %% special heading for the titlepage

%%% title
\title{Fréchet Distance}

%%% author(s)
\author{Tom Stein}

%%% date
\date{November 27, 2022}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% set heading and footer
%%%

%%% scrheadings default: 
%%%      footer - middle: page number
\pagestyle{scrheadings}

%%% user specific
%%% usage:
%%% \position[heading/footer for the titlepage]{heading/footer for the rest of the document}

%%% heading - left
% \ihead[]{}

%%% heading - center
% \chead[]{}

%%% heading - right
% \ohead[]{}

%%% footer - left
% \ifoot[]{}

%%% footer - center
% \cfoot[]{}

%%% footer - right
% \ofoot[]{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% begin document
%%%

\begin{document}

% \pagenumbering{roman} %% small roman page numbers

%%% include the title
% \thispagestyle{empty}  %% no header/footer (only) on this page
%  \maketitle

% Titlepage ---------------------------------------------------------
%
\pdfbookmark{Titelpage}{pdf:title}
\include{titlepage}

%%% start a new page and display the table of contents
% \newpage
% \tableofcontents

%%% start a new page and display the list of figures
% \newpage
% \listoffigures

%%% start a new page and display the list of tables
% \newpage
% \listoftables

%%% display the main document on a new page 
\newpage

% \pagenumbering{arabic} %% normal page numbers (include it, if roman was used above)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% begin main document
%%% structure: \section \subsection \subsubsection \paragraph \subparagraph
%%%

\newgeometry{left=3cm, right=4.5cm, top=4cm, bottom=4cm}

\section{Intro}
\subsection{Conventions}
Some Text \cite{DBLP:journals/ijcga/AltG95}

%%%
%%% end main document
%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\appendix  %% include it, if something (bibliography, index, ...) follows below

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%
%%% bibliography
%%%
%%% available styles: abbrv, acm, alpha, apalike, ieeetr, plain, siam, unsrt
%%%
\bibliographystyle{alpha}

%%% name of the bibliography file without .bib
%%% e.g.: literature.bib -> \bibliography{literature}
\bibliography{literature}

\end{document}
%%% }}}
%%% END OF FILE

### 2022-11-05 docs/paper.tex
removed old scrpage2
docs: bump
wip: introduction chapter

@@
 \usepackage{amsmath,amssymb,amstext}
-
-%%%
-%%% layout
-%%%
-
-%%% scrpage2: KOMA heading and footer
-%%% Note: if you don't use this package, please remove 
-%%%       \pagestyle{scrheadings} and corresponding settings
-%%%       below too.
-\usepackage[automark]{scrpage2}
 
@@
 %%%
-%%% set heading and footer
-%%%
-
-%%% scrheadings default: 
-%%%      footer - middle: page number
-\pagestyle{scrheadings}
-
-%%% user specific
-%%% usage:
-%%% \position[heading/footer for the titlepage]{heading/footer for the rest of the document}
-
-%%% heading - left
-% \ihead[]{}
-
-%%% heading - center
-% \chead[]{}
-
-%%% heading - right
-% \ohead[]{}
-
-%%% footer - left
-% \ifoot[]{}
-
-%%% footer - center
-% \cfoot[]{}
-
-%%% footer - right
-% \ofoot[]{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%
 %%% begin document
@@
 
-\section{Intro}
-\subsection{Conventions}
-Some Text \cite{DBLP:journals/ijcga/AltG95}
+\section*{Abstract}
+
+\section{Introduction}
+% motivation
+
+% two (polygonal) curves
+
+% Hausdorff distance
+The commonly known \textit{Hausdorff distance} can be understood 
+as the largest distance between any two points on the curves 
+and is calculated by \autoref{eq_hausdorff_distance} \cite{alt_computing_1995}.
+
+\begin{align}
+  \label{eq_hausdorff_distance}
+  \delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)
+\end{align}
+
+However, this metric does not take the course of the curves into account. 
+Two curves may be very dissimilar in the preception of a human 
+but still have a small Hausdorff-distance, see \autoref{fig_hausdorff_distance_bad_example}.
+
+\begin{figure}[ht]
+  % TODO: insert figure
+  \caption[Hausdorff distance of two dissimilar curves]{The two shown curves are dissimilar (in a common sense) depsite having a small Hausdorff distance.}
+  \label{fig_hausdorff_distance_bad_example}
+\end{figure}
+
+% Fréchet distance motivation
+
+% continuous Fréchet distance equation
+
+% Scope
+
+% Fréchet distance Variants 
+\section{Variants}
+Through the years many slightly differnt variants of the continuous Fréchet distance have evolved. 
+The continuous Fréchet distance takes into account every single point on the curves, 
+with interpolation in the case of polygonal curves, 
+without allowing to walk backwards 
+and without any obstacles in between the two curves.
+
+% TODO: Why allow it? Better algorithms?
+The \textit{weak Fréchet distance} slightly modified the continuous version by allowing 
+going backwards.
+
+% TODO: Is that mapping bijective? Add sources
+The \textit{discrete Fréchet distance} 
+This gives an upper bound on the continuous Fréchet distance % Source?
+and is often used as a approximation. % Source?
+% discrete Fréchet distance and contrast to hausdorff distance
+
+% (Geodesic Fréchet distance)
+
+% (Homotopic Fréchet distance)
+
+\section{Algorithm}
+% Original algorithm for the continuous Fréchet distance
+
+% Free Space Diagram
+
+% Runtime analysis (?)
+% Correctness proof (?)
+
+\section{Conclusion}
 

### 2022-11-15 manim_frechet_distance/presentation.py
basic point on line distance animation

@@ -36,3 +36,126 @@ class MinimalPresentationExample(Scene):
 
         self.next_section("E", PresentationSectionType.NORMAL)
         self.play(moving_dot.animate.move_to(dots[6]), rate_func=linear)
+
+
+class BraceAnnotation(Scene):
+    def construct(self):
+        
+        dot = Dot([-2, -1, 0])
+        dot2 = Dot([2, 1, 0])
+        line = Line(dot.get_center(), dot2.get_center()).set_color(ORANGE)
+        b1 = Brace(line)
+        self.add(line, dot, dot2, b1)
+
+        self.next_section("Horizontal distance", PresentationSectionType.NORMAL)
+        b1text = b1.get_text("Horizontal distance")
+        self.play(Write(b1text))
+        
+
+        self.next_section("Diagonal distance", PresentationSectionType.NORMAL)
+        b2 = Brace(line, direction=line.copy().rotate(PI / 2).get_unit_vector())
+        b2text = b2.get_tex("x-x_1")
+        self.play(Create(b2))
+        self.play(Write(b2text))
+
+
+class ArgMinExample(Scene):
+    def construct(self):
+        ax = Axes(
+            x_range=[0, 10], y_range=[0, 100, 10], axis_config={"include_tip": False}
+        )
+        labels = ax.get_axis_labels(x_label="x", y_label="f(x)")
+
+        t = ValueTracker(0)
+
+        def func(x):
+            return 2 * (x - 5) ** 2
+        graph = ax.plot(func, color=MAROON)
+
+        initial_point = [ax.coords_to_point(t.get_value(), func(t.get_value()))]
+        dot = Dot(point=initial_point)
+
+        dot.add_updater(lambda x: x.move_to(ax.c2p(t.get_value(), func(t.get_value()))))
+        x_space = np.linspace(*ax.x_range[:2],200)
+        minimum_index = func(x_space).argmin()
+
+        self.add(ax, labels)
+        self.next_section("Axes", PresentationSectionType.NORMAL)
+
+        self.next_section("Function")
+        self.play(Create(graph))
+
+        self.next_section("Point")
+        self.play(Create(dot))
+
+        self.next_section("Minimum", PresentationSectionType.NORMAL)
+        self.play(t.animate.set_value(x_space[minimum_index]))
+
+
+class FrechetDistanceExample(Scene):
+    def construct(self):
+        ax = Axes(
[…]

### 2022-11-15 manim_frechet_distance/presentation.py
fixed Fréchet distance example

@@ -95,7 +95,7 @@ class ArgMinExample(Scene):
 class FrechetDistanceExample(Scene):
     def construct(self):
         ax = Axes(
-            x_range=[0, 10], y_range=[0, 100, 10], axis_config={"include_tip": False}
+            x_range=[0, 10], y_range=[0, 100, 10], axis_config={"include_tip": False, "include_numbers": True}
         )
         labels = ax.get_axis_labels(x_label="x", y_label="y")
 
@@ -109,7 +109,7 @@ class FrechetDistanceExample(Scene):
         def func1(x):
             return 2 * (x - 5) ** 2
         def func2(x):
-            return -1.5 * (x - 1) ** 2 + 61.5
+            return -1.5 * (x - 2) ** 2 + 81
         graph1 = ax.plot(func1, color=MAROON)
         graph2 = ax.plot(func2, color=BLUE)
 
@@ -138,7 +138,7 @@ class FrechetDistanceExample(Scene):
 
         def dist_text():
             d = np.sqrt((x1.get_value() - x2.get_value())**2 + (func1(x1.get_value()) - func2(x2.get_value()))**2)
-            return Text(f"{d:.2f}", font_size=0.5 * DEFAULT_FONT_SIZE).next_to(line.get_center(), LEFT)
+            return Text(f"{d:.2f}", font_size=0.5 * DEFAULT_FONT_SIZE).next_to(line.get_center(), RIGHT)
         dist = dist_text()
         dist.add_updater(lambda t: t.become(dist_text()))
         self.play(Create(dist))
@@ -156,6 +156,18 @@ class FrechetDistanceExample(Scene):
         self.wait(1)
 
         self.next_section("Together to 0", PresentationSectionType.NORMAL)
-        self.play(x1.animate.set_value(0), x2.animate.set_value(0), run_time=3, rate_func=linear)
+        self.play(x1.animate.set_value(0), x2.animate.set_value(0), run_time=3, rate_func=smooth)
         self.wait(1)
+
+        self.next_section("Max vertical dist", PresentationSectionType.NORMAL)
+        self.play(x1.animate.set_value(26/7), x2.animate.set_value(26/7), run_time=2)
+        self.wait(1)
+        self.play(x1.animate.set_value(26/7+0.5), x2.animate.set_value(26/7+0.5))
+        self.wait(1)
+        self.play(x1.animate.set_value(26/7-0.5), x2.animate.set_value(26/7-0.5))
[…]

### 2022-11-16 manim_frechet_distance/presentation.py
change name to Fréchet-Distance in deployed version

@@ -162,10 +162,14 @@ class FrechetDistanceExample(Scene):
         self.next_section("Max vertical dist", PresentationSectionType.NORMAL)
         self.play(x1.animate.set_value(26/7), x2.animate.set_value(26/7), run_time=2)
         self.wait(1)
+
+        self.next_section("Max vertical dist right", PresentationSectionType.NORMAL)
         self.play(x1.animate.set_value(26/7+0.5), x2.animate.set_value(26/7+0.5))
-        self.wait(1)
+
+        self.next_section("Max vertical dist left", PresentationSectionType.NORMAL)
         self.play(x1.animate.set_value(26/7-0.5), x2.animate.set_value(26/7-0.5))
-        self.wait(1)
+        
+        self.next_section("Max vertical dist", PresentationSectionType.NORMAL)
         self.play(x1.animate.set_value(26/7), x2.animate.set_value(26/7))
         self.wait(1)
 

### 2022-11-16 manim_frechet_distance/presentation.py
replaced Text with DecimalNumber
increases rendering performance by a lot

@@ -127,6 +127,8 @@ class FrechetDistanceExample(Scene):
         initial_point2 = [ax.coords_to_point(x2.get_value(), func2(x2.get_value()))]
         dot2 = Dot(point=initial_point2)
         dot2.add_updater(lambda x: x.move_to(ax.c2p(x2.get_value(), func2(x2.get_value()))))
+        dot1.set_z_index(1)
+        dot2.set_z_index(1)
 
         self.next_section("Points")
         self.play(Create(dot1), Create(dot2))
@@ -136,12 +138,13 @@ class FrechetDistanceExample(Scene):
         line.add_updater(lambda z: z.become(Line(dot1.get_center(), dot2.get_center(), color=ORANGE)))
         self.play(Create(line))
 
-        def dist_text():
+        def dist():
             d = np.sqrt((x1.get_value() - x2.get_value())**2 + (func1(x1.get_value()) - func2(x2.get_value()))**2)
-            return Text(f"{d:.2f}", font_size=0.5 * DEFAULT_FONT_SIZE).next_to(line.get_center(), RIGHT)
-        dist = dist_text()
-        dist.add_updater(lambda t: t.become(dist_text()))
-        self.play(Create(dist))
+            return d
+        distNumber = DecimalNumber(dist()).next_to(line.get_center(), RIGHT)
+        distNumber.add_updater(lambda t: t.set_value(dist()))
+        distNumber.add_updater(lambda t: t.next_to(line.get_center(), RIGHT))
+        self.play(Create(distNumber))
         self.wait(1)
 
         x_space = np.linspace(*ax.x_range[:2],200)
@@ -171,6 +174,10 @@ class FrechetDistanceExample(Scene):
         
         self.next_section("Max vertical dist", PresentationSectionType.NORMAL)
         self.play(x1.animate.set_value(26/7), x2.animate.set_value(26/7))
+        self.wait(1)
+
+        self.next_section("Diagonal", PresentationSectionType.NORMAL)
+        self.play(x1.animate.set_value(26/7-0.1), x2.animate.set_value(26/7+0.1))
         self.wait(1)
 
 

### 2022-11-17 docs/paper.tex
basic point on line distance animation
wip: intro

@@
 % motivation
+At first it seems that measuring the similairty between two curves
+is a very theoretical problem
+but it has many, not that obvious, pratical applications.
+It is used in Geo-Information-Systems (GIS) to 
+quantify the similarity or quality of two maps, % cite
+find patterns in vehicle traffic or animal tracking data 
+by trajectory analysis, % cite
+gesture recognition on touch devices like smartphones, % cite
+and shape matching in the field of computer vision.
+
+To quantify the similairty of two given curves (e.g. shapes, paths, or trajectories)
+a distance metric can be used. 
+We can't simply use a distance metric like eucledian distance 
+becuase we don't want to compare points to points 
+but instead compare curves to curves.
 
 % two (polygonal) curves
+% TODO: replace continous funcitons with other name
+In the following we'll focus on polygonal curves (polygonal chains) 
+instead of continuous functions ofen used in pure mathematics 
+because most data we have for these problems is 
+captured by some none continuous mechanisms, e.g., GPS positions from a tracking device.
+Therefore we consider a polygonal curve $P$ to be definded as a 
+sequence of points $p_1, \dots, p_n \in R^d$ where $d$ is the dimensionality 
+of the space, e.g., $n=2$ or $n=3$, and linear interpolation between this points. 
+However, many concepts also apply to continuous functions.
 
@@
 However, this metric does not take the course of the curves into account. 
-Two curves may be very dissimilar in the preception of a human 
-but still have a small Hausdorff-distance, see \autoref{fig_hausdorff_distance_bad_example}.
+Two curves may have a small Hausdorff-distance but seem very dissimilar 
+when looking at them, see \autoref{fig_hausdorff_distance_bad_example}.
 
@@
 % TODO: Why allow it? Better algorithms?
-The \textit{weak Fréchet distance} slightly modified the continuous version by allowing 
+The \textit{weak Fréchet distance} slightly modifies the continuous version by allowing 
 going backwards.
@@
 
-\section{Algorithm}
-% Original algorithm for the continuous Fréchet distance
+\section{Algorithms}
+% Chapter summary and structure
 
+\subsection{Original algorithm for the continuous Fréchet distance}
 % Free Space Diagram
-
 % Runtime analysis (?)
 % Correctness proof (?)
+
+\subsection{Improved Version?}
+
+\subsection{Algorithms for different variants}
+% discrete Fréchet distance
+
+% weak Fréchet distance
 

### 2022-11-21 docs/paper.tex
Hausdorff distance graphics

@@
 The commonly known \textit{Hausdorff distance} can be understood 
-as the largest distance between any two points on the curves 
-and is calculated by \autoref{eq_hausdorff_distance} \cite{alt_computing_1995}.
+as the largest distance between any two points on the curves. 
+For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}.
+The Hausdorff distance is mathematically formulated as stated in \autoref{eq_hausdorff_distance} \cite{alt_computing_1995}.
 
@@
   % TODO: insert figure
-  \caption[Hausdorff distance of two dissimilar curves]{The two shown curves are dissimilar (in a common sense) depsite having a small Hausdorff distance.}
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+      }
+      \caption{Hausdorff distance between the red curve $P$ and green curve $Q$. }
+      \label{fig_hausdorff_distance_example}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
+      }
+      \caption[Hausdorff distance of two dissimilar curves]{Two dissimilar curves with a small Hausdorff distance.}
       \label{fig_hausdorff_distance_bad_example}
+  \end{subfigure}
+  \caption[Hausdorff distance examples]{Hausdorff distance examples.}
+  \label{fig_hausdorff_distance_examples}
 \end{figure}

### 2022-11-22 docs/paper.tex
add Fréchet distance to intro section

@@
 %%%
+\RequirePackage[table,dvipsnames]{xcolor}
+\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}
 
-\RequirePackage[table,dvipsnames]{xcolor}
-
-\definecolor{TUGreen}{rgb}{0.517,0.721,0.094}
+%%%
+%%% Macros
+%%%
+\newtheorem{mydef}{Definition}
 
@@
 by trajectory analysis, % cite
-gesture recognition on touch devices like smartphones, % cite
+gesture recognition on touch devices like smartphones \cite{hu_research_2022}, % cite
 and shape matching in the field of computer vision.
@@
 % TODO: replace continous funcitons with other name
-In the following we'll focus on polygonal curves (polygonal chains) 
-instead of continuous functions ofen used in pure mathematics 
-because most data we have for these problems is 
-captured by some none continuous mechanisms, e.g., GPS positions from a tracking device.
-Therefore we consider a polygonal curve $P$ to be definded as a 
-sequence of points $p_1, \dots, p_n \in R^d$ where $d$ is the dimensionality 
-of the space, e.g., $n=2$ or $n=3$, and linear interpolation between this points. 
-However, many concepts also apply to continuous functions.
+\begin{mydef}
+  \label{def_curve}
+  We define a curve as a continuous function $f : [a,b] \mapsto V$ 
+  with $a,b \in \mathbb{R}, a \leq b$ and $V$ being the space of the curve,
+  e.g., $\mathbb{R}^2$ or $\mathbb{R}^3$.
+\end{mydef}
+
+\begin{mydef}
+  \label{def_polygonal_curve}
+  We define a polygonal curve (polygonal chain) 
+  as a linearly interpolated curve 
+  on $n \in \mathbb{N}$ points in $V$ such that
+  $P : [0,n] \mapsto V$ 
+  with  $P(i + \lambda) = (1 - \lambda)P(i) + \lambda P(i+1)$
+  for all $\lambda \in [0,1]$ and all $i \in \{0, 1, \dots, n-1\}$.
+\end{mydef}
+
+In the following we will focus on polygonal curves (polygonal chains) 
+instead of classical continuous curves, ofen used in pure mathematics, 
+because most data we have in this area is 
+captured by some none continuous mechanism, e.g., GPS positions from a tracking device.
+Addionally, we focus on the 2D Space, i.e. $V = \mathbb{R}^2$.
+However, many concepts also apply to normal curves and higher dimensional spaces.
 
@@
 as the largest distance between any two points on the curves. 
-For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}.
-The Hausdorff distance is mathematically formulated as stated in \autoref{eq_hausdorff_distance} \cite{alt_computing_1995}.
+For each curve, a point with the maximum shortest distance to a point on the other curve is searched for, see \autoref{fig_hausdorff_distance_example}. \cite{alt_computing_1995}
 
-\begin{align}
-  \label{eq_hausdorff_distance}
-  \delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)
-\end{align}
+\begin{mydef}
+  \label{def_hausdorff_distance}
+  Given two (polygonal) curves $P$ and $Q$, 
+  their Hausdorff distance $\delta_{hd}(P,Q)$ is mathematically defined as: 
+  $$\delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)$$
+\end{mydef}
 
-However, this metric does not take the course of the curves into account. 
+This metric is very intuitive and can be easily computed for polygonal curves \cite{ko_complexity_2013}.
+It finds applications in computer vision and computer graphics 
+where it is used to compare the similarity of two mesh objects \cite{cignoni_metro_1998,ko_complexity_2013}. 
+However, this metric does not take the course of the curves into account, 
+but only the point sets of the individual curves.
 Two curves may have a small Hausdorff-distance but seem very dissimilar 
@@
 % Fréchet distance motivation
+In reality, the course of the curves is important for many applications. 
+For example, in handwriting character recognition on digital input devices, especially in Chinese, % TODO: add more citations
+gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
+and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
+For this reason, the \textit{Fréchet distance}, named after the famous french mathematician René Maurice Fréchet, was introduced.
+
+A common analogy for the Fréchet distance can be shown by a man walking his dog. 
+Assume that both the man and the dog are moving along a predefined fixed curve. 
+Both are free to adjust their speed as they wish, even to stop. 
+However, they are not allowed to go backwards. 
+We are looking for the minimum length of the dog leash 
+that is necessary for both of them to reach their destination. 
+The length of the dog leash is then exactly the Fréchet distance of the two curves.
 
@@
 
+\begin{mydef}
+  \label{def_frechet_distance}
+  Given two (polygonal) curves $P: [a,a'] \mapsto V$, $Q: [b,b'] \mapsto V$ and a distance function $\text{dist}: (V,V) \mapsto \mathbb{R}$, 
+  their Fréchet distance $\delta_{F}(P,Q)$ is mathematically defined as: 
+  $$\delta_{F}(P,Q) = \inf_{\substack{\alpha: [0,1] \mapsto [a, a'] \\ \beta: [0,1] \mapsto [b, b']}} \max_{t \in [0,1]} \text{dist}(P(\alpha(t)), Q(\beta(t)))$$
+  where $\alpha$ and $\beta$ represend the walking speed as 
+  continuous monotonically increasing functions with
+  $\alpha(0) = a, \alpha(1) = a'$, 
+  $\beta(0) =b, \beta(1) =b'$.
+\end{mydef}
+
+% TODO: Add example of Frechet distance
+
 % Scope
+% - Different Variants and their differences
+% - Original Algorithm by Alt and Godau
+% - Original Algorihm for discrete Fréchet distance
+% - Newer Algorithms (vor variants)
 
@@
 \section{Variants}
-Through the years many slightly differnt variants of the continuous Fréchet distance have evolved. 
+Through the years many slightly differnt variants of the continuous Fréchet distance (see Definition \ref{def_frechet_distance}) have evolved. 
 The continuous Fréchet distance takes into account every single point on the curves, 

### 2022-11-26 docs/paper.tex
wip: discrete frechet distance

@@
 %%%
-\RequirePackage[table,dvipsnames]{xcolor}
+\RequirePackage[table, dvipsnames]{xcolor}
 \definecolor{TUGreen}{rgb}{0.517,0.721,0.094}
@@
 \newtheorem{mydef}{Definition}
+\newcommand{\mydefautorefname}{Definition}
 
@@
 % motivation
-At first it seems that measuring the similairty between two curves
-is a very theoretical problem
-but it has many, not that obvious, pratical applications.
+At first, it seems that measuring the similarity between two curves
+is a very theoretical problem, but it has many, not that obvious, practical applications.
 It is used in Geo-Information-Systems (GIS) to 
@@
 
-To quantify the similairty of two given curves (e.g. shapes, paths, or trajectories)
+To quantify the similarity of two given curves (e.g. shapes, paths, or trajectories)
 a distance metric can be used. 
-We can't simply use a distance metric like eucledian distance 
-becuase we don't want to compare points to points 
+We can't simply use a distance metric like euclidean distance 
+because we don't want to compare points to points 
 but instead compare curves to curves.
@@
   as a linearly interpolated curve 
-  on $n \in \mathbb{N}$ points in $V$ such that
-  $P : [0,n] \mapsto V$ 
-  with  $P(i + \lambda) = (1 - \lambda)P(i) + \lambda P(i+1)$
-  for all $\lambda \in [0,1]$ and all $i \in \{0, 1, \dots, n-1\}$.
+  on $n \in \mathbb{N}$ points in $V$.
+  Therefore, $P : [0,n] \mapsto V$ 
+  should be affine on the interval $[i,i+1]$ for all $i \in \{0, 1, \dots, n-1\}$, 
+  i.e., $P(i + \lambda) = (1 - \lambda)P(i) + \lambda P(i+1)$ for all $\lambda \in [0,1]$.
 \end{mydef}
@@
 In the following we will focus on polygonal curves (polygonal chains) 
-instead of classical continuous curves, ofen used in pure mathematics, 
+as defined in \autoref{def_polygonal_curve} 
+instead of classical curves (\autoref{def_curve}), often used in pure mathematics, 
 because most data we have in this area is 
-captured by some none continuous mechanism, e.g., GPS positions from a tracking device.
-Addionally, we focus on the 2D Space, i.e. $V = \mathbb{R}^2$.
+captured by some noncontinuous mechanism, e.g., GPS positions from a tracking device.
+Additionally, we focus on the 2D Space, i.e. $V = \mathbb{R}^2$.
 However, many concepts also apply to normal curves and higher dimensional spaces.
@@
 
-This metric is very intuitive and can be easily computed for polygonal curves \cite{ko_complexity_2013}.
+The Hausdorff distance metric $\delta_{hd}(P,Q)$ defined in \autoref{def_hausdorff_distance} 
+is very intuitive and can be easily computed for polygonal curves \cite{ko_complexity_2013}.
 It finds applications in computer vision and computer graphics 
@@
 \begin{figure}[ht]
-  % TODO: insert figure
   \centering
@@
 and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
-For this reason, the \textit{Fréchet distance}, named after the famous french mathematician René Maurice Fréchet, was introduced.
+For this reason, the \textit{Fréchet distance}, named after the famous French mathematician René Maurice Fréchet, was introduced.
 
@@
 % continuous Fréchet distance equation
-
 \begin{mydef}
@@
   $$\delta_{F}(P,Q) = \inf_{\substack{\alpha: [0,1] \mapsto [a, a'] \\ \beta: [0,1] \mapsto [b, b']}} \max_{t \in [0,1]} \text{dist}(P(\alpha(t)), Q(\beta(t)))$$
-  where $\alpha$ and $\beta$ represend the walking speed as 
+  where $\alpha$ and $\beta$ represent the walking speed as 
   continuous monotonically increasing functions with
@@
 
-% TODO: Add example of Frechet distance
+% TODO: Add example graphic of Frechet distance
+
+Besides the previously mentioned applications, 
+where the course of the curves is important, 
+the Fréchet distance also finds applications in
+(online) signature verification \cite{zheng_algorithm_2008,fang_research_2018},
+map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
+and moving object analysis, e.g., detecting commuting patterns \cite{buchin_detecting_2011}.
 
@@
 \section{Variants}
-Through the years many slightly differnt variants of the continuous Fréchet distance (see Definition \ref{def_frechet_distance}) have evolved. 
-The continuous Fréchet distance takes into account every single point on the curves, 
-with interpolation in the case of polygonal curves, 
+Through the years many slightly different variants of the continuous Fréchet distance (see Definition \ref{def_frechet_distance}) have evolved. 
+The continuous Fréchet distance takes into account every single point on the curves 
 without allowing to walk backwards 
@@
 
-% TODO: Is that mapping bijective? Add sources
-The \textit{discrete Fréchet distance} 
-This gives an upper bound on the continuous Fréchet distance % Source?
-and is often used as a approximation. % Source?
+% Discrete Fréchet distance
+The \textit{discrete Fréchet distance} is specifically tied to 
+polygonal curves (\autoref{def_polygonal_curve}) 
+because it requires a finite number of vertices.
+The dog walking analogy can be adapted by replacing both the man and the dog by two frogs,
+which are only allowed to jump from their current vertex to their next 
+vertex in forward direction \cite{bringmann_why_2014}.
+While it is allowed for both frogs to jump simultaneously,
+it is not allowed to skip vertices while jumping. 
+We ensure this behavior by defining a coupling between the curves in \autoref{def_point_coupling}.
+
+\begin{mydef}
+  \label{def_point_coupling}
+  A \textit{coupling} L between two polygonal curves $P$ and $Q$, 
+  with points $(u_1, \dots, u_p)$ and $(v_1, \dots, v_p)$ respectively, 
+  is defined as a sequence
+  $$(u_{a_1}, v_{b_1}),(u_{a_2}, v_{b_2}), \dots, (u_{a_m}, v_{b_m})$$
+  such that $a_1 = 1, b_1 = 1, a_m = p, b_m = q$ 
+  while preserving order and the jump length limitation.
+  % Note different definiton than in original paper
+  % fixes error for unequal length polygonal curves
+  Thus, for all $i \in \{1, \dots m-1\}$
+  the jump length should be positive but no more than 1,
+  i.e., $0 \leq a_{i+1} - a_i \leq 1$ (respectively $0 \leq b_{i+1} - b_i \leq 1$).
+\end{mydef}
+
+Note that $P$ and $Q$ may be different in terms of number of points,
+implying that the frog on the shorter curve 
+has to stop and wait on at least one vertex once
+due to the pigeonhole principle.
+To order the possible couplings between $P$ and $Q$ we define the 
+\textit{length} of a coupling in \autoref{def_coupling_length}.
+
+\begin{mydef}
+  \label{def_coupling_length}
+  The \textit{length} $||L||$ of a coupling $L$ is the greatest distance of the $m$ pairs in $L$, that is, 
+  $$||L|| = \max_{i=1, \dots, m} dist(u_{a_i}, v_{b_i})$$
+\end{mydef}
+
+The definition of the discrete Fréchet distance immediately follows from 
+the length of a coupling and is given by \autoref{def_discrete_frechet_distance}.
+
+\begin{mydef}
+  \label{def_discrete_frechet_distance}
+  Given two polygonal curves $P$ and $Q$
+  their discrete Fréchet distance $\delta_{dF}(P,Q)$ is defined as the shortest coupling between them: 
+  $$\delta_{dF}(P,Q) = \min \left\{ ||L|| \ | \  L \text{ is a coupling between P and Q} \right\}$$
+\end{mydef}
+
+The discrete variation (\autoref{def_discrete_frechet_distance}) gives an upper bound on the continuous Fréchet distance (\autoref{def_frechet_distance}). 
+% Clearify if this is really an upper bound for polygonal curves with example
+This might not be very obvious for the case of polygonal curves, 
+but \autoref{fig_frechet_distance_upper_bound_example} gives an intuitive example.
+
+\begin{figure}[ht]
+  \centering
+  % TODO: Replace graphic with appropiate example
+  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \caption[Upper bound on Fréchet distance]{
+    In the case of discrete Fréchet distance 
+    one frog has to wait at either start or end of his path 
+    while the other one is at the middle vertex, 
+    thus requiring a minimal length of $\sqrt{d^2 + l^2}$ for any coupling.
+    In contrast, a leash of length $d$ is sufficient for the continuous Fréchet distance.}
+  \label{fig_frechet_distance_upper_bound_example}
+\end{figure}
+
+Additionally, it's also a lower bound on the continuous Fréchet distance 
+plus the distance of the longest legal jump of any frog. 
+By increasing the number of points of a polygonal curve that 
+approximates a polygonal curve 
+and can be used as an approximation. % Source?
+\cite{eiter_computing_nodate}
+
 % discrete Fréchet distance and contrast to hausdorff distance
+% The discrete Fréchet distance also assings each point another point on the other curve
+% just like the hausdorff distance, however it additionally preserves the order of points.
 

### 2022-11-28 docs/paper.tex
wip: original algorithm

@@
 
-\newgeometry{left=3cm, right=4.5cm, top=4cm, bottom=4cm}
+\newgeometry{left=3cm, right=4cm, top=3cm, bottom=3cm}
 
@@
 The length of the dog leash is then exactly the Fréchet distance of the two curves.
+A formal definition is given in \autoref{def_frechet_distance}.
 
@@
 % TODO: Why allow it? Better algorithms?
+\subsection{Weak Fréchet distance}
 The \textit{weak Fréchet distance} slightly modifies the continuous version by allowing 
@@
 % Discrete Fréchet distance
+\subsection{Discrete Fréchet distance}
 The \textit{discrete Fréchet distance} is specifically tied to 
@@
 % Chapter summary and structure
+Several algorithms to compute the Fréchet distance of polygonal curves were proposed, 
+starting from the Original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
+The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions. 
+This chapter is structured as follows: 
+firstly the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
+then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
+Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_improved_algorithm}.
+Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
+% Finally
 
 \subsection{Original algorithm for the continuous Fréchet distance}
+\label{sec_original_algorithm}
+The original algorithm to compute the continuous Fréchet distance (\autoref{def_frechet_distance})
+is based on the idea of solving a simpler decision problem and then doing a parameter search to find the exact value of $\delta_{F}$.
+The decision problem is to decide whether two polygonal curves $P$ and $Q$ have a 
+Fréchet distance $\delta_{F}(P,Q) \leq \varepsilon$ for some fixed $\varepsilon \geq 0$.
+
+We start by looking at the simplest case for two polygonal curves with two vertices each, i.e., two line segments.
+The free space is are all valid combinations of positions on $P$ and $Q$ that have a distance of at most $\varepsilon$.
+Therefore, we define $F_\varepsilon = \{(s,t) \in [0,1]^2 \ | \ dist(P(s), Q(t)) \leq \varepsilon \}$.
+Plotting $s$ and $t$ on a two-dimensional graph, while indicating all illegal combinations in 
+another color, gives the free space diagram, see \autoref{fig_free_space}. \cite{alt_computing_1995}
+
 % Free Space Diagram
+\begin{figure}[ht]
+  \centering
+  % TODO: update figure graphics
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+      }
+      \caption{Line segments $P$ and $Q$}
+      \label{fig_free_space_lines_example}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
+      }
+      \caption{Free space diagram}
+      \label{fig_free_space_diagram_example}
+  \end{subfigure}
+  \caption[Free space diagram example]{
+    The free space diagram \autoref{fig_free_space_lines_example} 
+    for line segments $P$ and $Q$ and distance $\varepsilon$
+    from \autoref{fig_free_space_lines_example} using euclidean distance.}
+  \label{fig_free_space}
+\end{figure}
+
+The free space $F_\varepsilon$ has different shapes depending on the employed distance metric,
+e.g.,  the intersection of the unit square with an ellipse for euclidean distance 
+and with a parallelogram for $L_1$ or $L_\infty$. 
+The exact structure of the free space is not further relevant, 
+because only the convexity of the area is important. \cite{alt_computing_1995}
+
+We can now extend the free space diagram to longer polygonal curves 
+as stated in \autoref{def_free_space}.
+An example for a free space diagram on two non-trivial polygonal curves is given in \autoref{fig_complex_free_space}.
+
+\begin{mydef}
+  \label{def_free_space}
+  Given two polygonal curves $P$ and $Q$ with $p,q \in \mathbb{N}$ vertices respectively, 
+  the free space is defined as:
+  $$F_\varepsilon = \left\{ (s,t) \in [0,p] \times [0,q] \ | \ dist(P(s), Q(t)) \leq \varepsilon \right\}$$
+\end{mydef}
+
+\begin{figure}[ht]
+  \centering
+  % TODO: update figure graphics
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+      }
+      \caption{Polygonal curves $P$ and $Q$}
+      \label{fig_complex_free_space_lines_example}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.45\textwidth}
+      \resizebox{\textwidth}{!}{
+        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
+      }
+      \caption{Free space diagram}
+      \label{fig_complex_free_space_diagram_example}
+  \end{subfigure}
+  \caption[Complex free space diagram example]{
+    The free space diagram \autoref{fig_complex_free_space_diagram_example} 
+    for polygonal curves $P$ and $Q$ and distance $\varepsilon$ 
+    from \autoref{fig_complex_free_space_lines_example} using euclidean distance.}
+  \label{fig_complex_free_space}
+\end{figure}
+
+We observe that there exists 
+a monotonically increasing curve from $(0,0)$ to $(p,q)$ in $F_\varepsilon$
+if and only if the Fréchet distance between $P$ and $Q$ is less than or equal to $\varepsilon$,
+i.e., $\delta_F(P,Q) \leq \varepsilon$.
+Given such a curve, the walking speed functions $\alpha$ and $\beta$ from \autoref{def_frechet_distance}
+can be easily obtained by rescaling the parameter ranges.
+
+Computing the complete free space is very compute intensive 
+because each point in the continuous space needs to be computed. 
+Using the convexity property of each cell in the free space,
+we can make another observation requiring only a few points to be computed per cell. 
+These points define the legal regions on each boundary side of the cell, see \autoref{fig_legal_boundary_intervals_of_a_cell}. 
+We don't need to know the exact shape of the free space 
+because we travel in a monotonically increasing way through a convex space,
+hence never entering any illegal regions.
+
+\begin{figure}[ht]
+  \centering
+  % TODO: Replace graphic with appropiate example
+  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \caption[Legal intervals on free space cell boundary]{Legal intervals on the boundaries of a free space cell.}
+  \label{fig_legal_boundary_intervals_of_a_cell}
+\end{figure}
+
 % Runtime analysis (?)
-% Correctness proof (?)
+% Correctness proof (?)6
 
 \subsection{Improved Version?}
+\label{sec_improved_algorithm}
 
-\subsection{Algorithms for different variants}
-% discrete Fréchet distance
+% Algorithms for different variants
+\subsection{Algorithm for weak Fréchet distance}
 
-% weak Fréchet distance
+\subsection{Algorithm for discrete Fréchet distance}
+\label{sec_discrete_algorithm}
 
 \section{Conclusion}
+% Which version to use when
 

### 2022-11-30 docs/paper.tex
finish text of original algorithm

@@
 \subsection{Weak Fréchet distance}
-The \textit{weak Fréchet distance} slightly modifies the continuous version by allowing 
+The \textit{weak Fréchet distance} (non-monotone Fréchet distance) slightly modifies the continuous version by allowing 
 going backwards.
@@
 
-We observe that there exists 
-a monotonically increasing curve from $(0,0)$ to $(p,q)$ in $F_\varepsilon$
-if and only if the Fréchet distance between $P$ and $Q$ is less than or equal to $\varepsilon$,
-i.e., $\delta_F(P,Q) \leq \varepsilon$.
-Given such a curve, the walking speed functions $\alpha$ and $\beta$ from \autoref{def_frechet_distance}
-can be easily obtained by rescaling the parameter ranges.
-
-Computing the complete free space is very compute intensive 
+Computing the complete free space $F_\varepsilon$ is very compute intensive 
 because each point in the continuous space needs to be computed. 
 Using the convexity property of each cell in the free space,
-we can make another observation requiring only a few points to be computed per cell. 
+we can observe that only a small constant number of points need to be computed per cell. 
 These points define the legal regions on each boundary side of the cell, see \autoref{fig_legal_boundary_intervals_of_a_cell}. 
@@
   \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
-  \caption[Legal intervals on free space cell boundary]{Legal intervals on the boundaries of a free space cell.}
+  \caption[Legal intervals on free space cell boundary]{
+    Legal intervals on the boundaries of a free space cell. 
+    There are at most 8 intersection points for an ellipse and a rectangle.}
   \label{fig_legal_boundary_intervals_of_a_cell}
 \end{figure}
+
+Due to the constant number of intersection points required per cell, 
+the free space of a single cell can be computed in $\mathcal{O}(1)$ time.
+Therefore, the complete free space of all cells can be computed in $\mathcal{O}(pq)$
+because there are $p-1$ (respectively $q-1$) edges in $P$ (respectively $Q$).
+
+We observe that there exists 
+a monotonically increasing curve from $(0,0)$ to $(p,q)$ in $F_\varepsilon$
+if and only if the Fréchet distance between $P$ and $Q$ is less than or equal to $\varepsilon$,
+i.e., $\delta_F(P,Q) \leq \varepsilon$.
+Given such a curve, the walking speed functions $\alpha$ and $\beta$ from \autoref{def_frechet_distance}
+can be easily obtained by rescaling the parameter ranges. 
+Computing such a monotonically increasing curve can be done using dynamic programming 
+by computing the bottom left most reachable points in each cell starting at $(0,0)$. 
+In each step the solution for a single cell can be computed in constant time based on the 
+solutions to previously computed cells. 
+Therefore, the runtime is also $\mathcal{O}(pq)$. 
+See \autoref{fig_free_space_curve} for an example curve. \cite{alt_computing_1995}
+
+\begin{figure}[ht]
+  \centering
+  % TODO: Replace graphic with appropiate example
+  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \caption{The monotonically increasing curve through the free space is computed by dynamic programming.}
+  \label{fig_free_space_curve}
+\end{figure}
+
+% Actually computing the exact Fréchet distance
+So far we discussed the decision problem whether $\delta_{F}(P,Q) \leq \varepsilon$
+and observe that it is solvable in $\mathcal{O}(pq)$ time.
+What we actually want to know is the exact value of the Fréchet distance.
+A straight forward approach for this is to compute the value bit by bit 
+using binary search in total $\mathcal{O}(pq \log (\text{``accuracy bits''}))$ time.
+In practice this approach works really well 
+and is often favored over other more complicated approaches. \cite{alt_computing_1995}
+
+Another approach is to check all values of $\varepsilon$ where something ``interesting'' happens
+until the smallest value of $\varepsilon$ is obtained where a monotonically increasing path 
+from $(0,0)$ to $(p,q)$ in $F_\varepsilon$ exists.
+We will call these the \textit{critical values} of $\varepsilon$, which are:
+\begin{enumerate}
+  \item the minimal value of $\varepsilon$ where $(0,0) \in F_\varepsilon$ and $(p,q) \in F_\varepsilon$.
+  \item values of $\varepsilon$ where a new passage on the border edge 
+        between two neighboring cells opens up.
+  \item values of $\varepsilon$ where a new passage horizontal (respectively vertical) passage 
+        between two non-neighboring cells opens up.
+\end{enumerate}
+
+There is one value of case 1, 
+specifically the maximum of starting point to starting point and endpoint to endpoint distance,
+$\mathcal{O}(pq)$ values of case 2, 
+namely the number of vertex to edge distances,
+and $\mathcal{O}(p^2q + pq^2)$ values of case 3,
+particularly the number of times two vertices of one curve have the same distance to one edge of the other curve.
+Determining all these values can be done in $\mathcal{O}(p^2q + pq^2)$ time 
+because each value in each case can be computed in constant time, 
+making the values of the third case dominate the time complexity.
+% TODO: why does this simplyfication work
+These values need to be sorted in $\mathcal{O}((p^2q + pq^2) \log (p^2q + pq^2)) = \mathcal{O}((p^2q + pq^2) \log (pq))$ time
+in order to perform binary search in $\mathcal{O}(pq \log (p^2q + pq^2))$ time on them
+using the algorithm for the decision problem.
+The overall time required for this approach is dominated by the sorting step, 
+thus $\mathcal{O}((p^2q + pq^2) \log (pq))$.
+This time complexity can be decreased to $\mathcal{O}(pq \log (pq))$ 
+using parametric search \cite{megiddo_applying_1983, cole_slowing_1987} to work faster in theory.
+However, due to the specific internally used sorting network 
+it is hard to implement this efficiently in practice. \cite{alt_computing_1995}
 

### 2022-11-30 docs/paper.tex
cleanup

@@
 polygonal curves (\autoref{def_polygonal_curve}) 
-because it requires a finite number of vertices.
+because it requires a finite number of vertices 
+and forbids walking or staying in between them.
 The dog walking analogy can be adapted by replacing both the man and the dog by two frogs,
@@
 By increasing the number of points of a polygonal curve that 
-approximates a polygonal curve 
-and can be used as an approximation. % Source?
-\cite{eiter_computing_nodate}
+over approximates some other polygonal curve
+this property can be used to compute the continuous Fréchet distance \cite{eiter_computing_1994}.
 

### 2022-12-01 docs/paper.tex
wip: Fréchet distance free space figures

@@
 % TODO: Why allow it? Better algorithms?
-\subsection{Weak Fréchet distance}
+\subsection{Weak Fréchet Distance}
 The \textit{weak Fréchet distance} (non-monotone Fréchet distance) slightly modifies the continuous version by allowing 
@@
 % Discrete Fréchet distance
-\subsection{Discrete Fréchet distance}
+\subsection{Discrete Fréchet Distance}
 The \textit{discrete Fréchet distance} is specifically tied to 
@@
 
+\subsection{Fréchet Distance for Uncertain Curves}
+
 \section{Algorithms}
@@
 
-\subsection{Original algorithm for the continuous Fréchet distance}
+\subsection{Original Algorithm for the Continuous Fréchet Distance}
 \label{sec_original_algorithm}
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+        \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
       }
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
+        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
       }
@@
   \caption[Free space diagram example]{
-    The free space diagram \autoref{fig_free_space_lines_example} 
+    The free space diagram \autoref{fig_free_space_diagram_example} 
     for line segments $P$ and $Q$ and distance $\varepsilon$
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+        \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
       }
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-bad-example.pdf}
+        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
       }
@@
     The free space diagram \autoref{fig_complex_free_space_diagram_example} 
-    for polygonal curves $P$ and $Q$ and distance $\varepsilon$ 
-    from \autoref{fig_complex_free_space_lines_example} using euclidean distance.}
+    for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
+    and distance $\varepsilon = 0.73$ using euclidean distance.}
   \label{fig_complex_free_space}

### 2022-12-01 docs/paper.tex
add description and algorithm for weak Fréchet distance

@@
 and without any obstacles in between the two curves.
+In this chapter the most prominent variants of the Fréchet distance are shortly described 
+while algorithms for these are given in the following \autoref{sec_algorithms}.
 
@@
 \subsection{Weak Fréchet Distance}
-The \textit{weak Fréchet distance} (non-monotone Fréchet distance) slightly modifies the continuous version by allowing 
-going backwards.
+\label{sec_weak_frechet_distance}
+The \textit{weak Fréchet distance} (non-monotone Fréchet distance) 
+slightly modifies the continuous version by allowing going backwards. 
+The definition remains the same as for the continuous Fréchet distance in \autoref{def_frechet_distance}
+where the monotonicity of $\alpha$ and $\beta$ is no longer required.
+Note that they are still required to be continuous and have the correct starting and ending values.
+% TODO: Add use cases
+The weak Fréchet distance is a lower bound on the continuous Fréchet distance.
+However, the Ratio between them may be arbitrarily large, 
+as can be seen by the structure given in \autoref{fig_weak_frechet_distance}. \cite{alt_computing_1995}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.8\textwidth]{images/frechet_distance/weak-frechet-distance-example.pdf}
+  \caption{
+    The two shown polygonal curves $P$ and $Q$ have a small weak Fréchet distance 
+    because the point on $Q$ can be moved forward and backward as often as needed. 
+    The continuous Fréchet distance is rather large 
+    because the point on $Q$ can only move forward.}
+  \label{fig_weak_frechet_distance}
+\end{figure}
 
@@
   % TODO: Replace graphic with appropiate example
-  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \includegraphics[width=0.8\textwidth]{images/frechet_distance/discrete-frechet-distance-upper-bound.pdf}
   \caption[Upper bound on Fréchet distance]{
-    In the case of discrete Fréchet distance 
-    one frog has to wait at either start or end of his path 
-    while the other one is at the middle vertex, 
+    In the case of the discrete Fréchet distance 
+    one frog has to wait at either start or end $Q$
+    while the other one is at the middle vertex of $P$, 
     thus requiring a minimal length of $\sqrt{d^2 + l^2}$ for any coupling.
@@
 
+\subsection{Fréchet distance with speed limits}
+
 \subsection{Fréchet Distance for Uncertain Curves}
@@
 \section{Algorithms}
+\label{sec_algorithms}
 % Chapter summary and structure
@@
 starting from the Original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
-The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions. 
+The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions, 
+e.g., the version by Buchin et al. \cite{buchin_four_2017}. 
 This chapter is structured as follows: 
-firstly the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
+first the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
 then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
@@
 A straight forward approach for this is to compute the value bit by bit 
-using binary search in total $\mathcal{O}(pq \log (\text{``accuracy bits''}))$ time.
+using binary search in total $\mathcal{O}(pq \log (\text{``accuracy''})) = \mathcal{O}(pq  (\text{``accuracy bits''}))$ time.
 In practice this approach works really well 
@@
 \subsection{Algorithm for weak Fréchet distance}
+\label{sec_weak_algorithm}
+The weak Fréchet distance, as described in \autoref{sec_weak_frechet_distance},
+the computation becomes less complex compared to the continuous Fréchet distance.
+Specifically, we can ignore critical values of the third type, 
+since it is no longer required to move monotonically 
+and one could just go left or down in the free space diagram taking any arbitrary continuous path.
+Therefore, the focus shifts towards critical values of the second type, 
+those where new passages between neighboring cells open up.
+Alt and Godau \cite{alt_computing_1995} proposed two algorithms to compute the weak Fréchet distance.
+
+For both algorithms the problem is modeled as a grid based graph of $p$ by $q$ vertices, 
+one for each cell of the free space diagram,
+and edges connecting the vertices with their, up to six, neighboring vertices.
+Additionally, $s$ is added as start vertex and $t$ as end vertex.
+This results in an undirected Graph $G=(V,E)$ 
+where $V = \{C_{i,j} \ | \ 0 \leq i \leq p, 0 \leq j \leq q \} \cup \{s,t\}$
+and $E=\{(C_{i,j}, C_{i+1,j}) \ | \ \ 0 \leq i < p, 0 \leq j \leq q\} \cup \{(C_{i,j}, C_{i,j+1}) \ | \ \ 0 \leq i \leq p, 0 \leq j < q\} \cup \{(s, C_{0,0}), (C_{p,q}, t)\}$.
+Each edge $(C_{i,j}, C_{i',j'})$ is assigned a critical minimal value of $\varepsilon$ 
+such that there is a passage through the border between two cells in the free space. 
+Furthermore, the minimal value of $\varepsilon$ such that $(0,0) \in F_\varepsilon$ is assigned to the first edge $(s, C_{0,0})$ 
+and the minimal value of $\varepsilon$ such that $(p,q) \in F_\varepsilon$ is assigned to the last edge $(C_{p,q}, t)$. \cite{alt_computing_1995}
+
+The first algorithm is again based on first solving the decision problem 
+and then searching for the right $\varepsilon$ through binary search.
+The decision problem is satisfied, if and only if there exists a path from $s$ to $t$ in $G$ 
+with a maximum single edge value of $\varepsilon$.
+This can be checked by removing all edged with values greater than $\varepsilon$ from $E$ 
+and then checking if $s$ is still connected to $t$. 
+This can be done through some graph traversal algorithm like breadth-first search or depth-first search
+in $\mathcal{O}(pq)$ time. \cite{alt_computing_1995}
+Since the number of critical values is no longer dominated by the third case,
+rather by the second case of $\mathcal{O}(pq)$ values, 
+the overall computation problem can be solved in $\mathcal{O}(pq \log (pq))$ time using binary search \cite{van_leusden_novel_2013}.
+
+The second algorithm works by directly searching for a path from $s$ to $t$ in $G$,
+without any prior removal of edges from $E$.
+This can be done using Prim's minimum spanning tree algorithm \cite{prim_shortest_1957} starting at $s$ 
+and stopping as soon as $t$ is included in the spanning tree. 
+The value of $\varepsilon$ is the value of the last added edge.
+Similarly, one can use Dijkstra's algorithm \cite{dijkstra_note_1959}, 
+optimizing for the minimum maximum edge value.
+Both approaches result in a time complexity of $\mathcal{O}(pq \log (pq))$. \cite{alt_computing_1995}
+In this case of a planar grid based graph with non-negative edge values 
+an even faster algorithm from Henzinger et al. \cite{henzinger_faster_1997} to solve the single source shortest paths 
+can be applied. This gives a time complexity of $\mathcal{O}(pq)$ to compute the weak Fréchet distance \cite{van_leusden_novel_2013}.
 

### 2022-12-01 docs/paper.tex
Algorithm for discrete Fréchet distance

@@
 then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
-Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_improved_algorithm}.
+Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_weak_algorithm}.
 Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
@@
 \label{sec_improved_algorithm}
+\cite{van_leusden_novel_2013}
+\cite{buchin_four_2017}
 
 % Algorithms for different variants
-\subsection{Algorithm for weak Fréchet distance}
+\subsection{Algorithm for the Weak Fréchet Distance}
 \label{sec_weak_algorithm}
@@
 
-\subsection{Algorithm for discrete Fréchet distance}
+\subsection{Algorithm for the Discrete Fréchet Distance}
 \label{sec_discrete_algorithm}
+The computation of the discrete Fréchet distance, as defined in \autoref{def_discrete_frechet_distance}, is very easy and straight forward 
+using a dynamic programming approach as originally suggested by Eiter et al. \cite{eiter_computing_1994}.
+Recall that the polygonal curves $P = (u_1, \dots, u_p)$ and $Q = (v_1, \dots, v_p)$ are defined as sequences of points.
+The algorithm works by recursively computing the minimal required $\varepsilon$ for partial curves 
+$P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
+using $c(i,j)$ as defined in \autoref{def_recursive_forumla_discrete_frechet_distance}.
+Starting from $c(1,1)$, which is the distance between the starting points $u_1$ and $v_1$,
+all other values of $c(i,j)$ are computed iteratively in a bottom up manner 
+and memoization on the previously computed values of $c(i,j)$.
+The resulting algorithm has runtime of $\mathcal{O}(pq)$
+because the constant time function $c(i,j)$ is called $pq$ times \cite{eiter_computing_1994}.
+
+\begin{mydef}
+  \label{def_recursive_forumla_discrete_frechet_distance}
+  The discrete Fréchet distance of polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
+  can be computed by
+  \begin{align*}
+    c(i,j) &= \begin{cases}
+      dist(u_1, v_1), &\text{if } i=1 \text{and } j=1 \\
+      \max \{c(i-1,j), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
+      \max \{c(i,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
+      \max \{c(i-1,j), c(i,j-1), c(i-1,j-1), dist(u_i, v_j)\}, &\text{if } i>1 \text{and } j>1
+    \end{cases} \\
+    &= \delta_{dF}(P',Q')
+  \end{align*}
+\end{mydef}
 

### 2022-12-02 docs/paper.tex
generate graphics as latex pgf plot

@@
 
+%%% pgf plots from matplotlib tikzplotlib
+\usepackage{pgfplots}
+\DeclareUnicodeCharacter{2212}{−}
+\usepgfplotslibrary{groupplots,dateplot}
+\usetikzlibrary{patterns,shapes.arrows}
+\pgfplotsset{compat=newest}
+
 %%% amsmath, amssymb, amstext: support for mathematics
@@
 %%%
-\RequirePackage[table, dvipsnames]{xcolor}
 \definecolor{TUGreen}{rgb}{0.517,0.721,0.094}
@@
 
-%%%
-%%% PDF
-%%%
-
-\usepackage{ifpdf}
-
-%%% Should be LAST usepackage-call!
-%%% For docu on that, see reference on package ``hyperref''
-\ifpdf%   (definitions for using pdflatex instead of latex)
-
-  %%% graphicx: support for graphics
-  \usepackage[pdftex]{graphicx}
-
-  \pdfcompresslevel=9
+\usepackage{graphicx}
 
@@
 ]{hyperref}
-
-  %%% provide all graphics (also) in this format, so you don't have
-  %%% to add the file extensions to the \includegraphics-command
-  %%% and/or you don't have to distinguish between generating
-  %%% dvi/ps (through latex) and pdf (through pdflatex)
-  \DeclareGraphicsExtensions{.pdf}
-
-\else %else   (definitions for using latex instead of pdflatex)
-
-  \usepackage[dvips]{graphicx}
-
-  \DeclareGraphicsExtensions{.eps}
-
-  \usepackage[%
-    dvips,           %% sets up hyperref for use with the dvips driver
-    colorlinks=false %% better for printout version; almost every hyperref-extension is eliminated by using dvips
-  ]{hyperref}
-
-\fi
 
@@
 the Fréchet distance also finds applications in
-(online) signature verification \cite{zheng_algorithm_2008,fang_research_2018},
+(online) handwriting signature verification \cite{zheng_algorithm_2008,fang_research_2018},
 map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
@@
 
-% TODO: Why allow it? Better algorithms?
 \subsection{Weak Fréchet Distance}
@@
   \centering
-  % TODO: Replace graphic with appropiate example
   \includegraphics[width=0.8\textwidth]{images/frechet_distance/discrete-frechet-distance-upper-bound.pdf}
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
+        % \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
+        \input{images/frechet_distance/line_segments.tex}
       }
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
+        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
+        \input{images/frechet_distance/free_space_single_cell.tex}
       }
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
+        % \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
+        \input{images/frechet_distance/polygonal_curves.tex}
       }
@@
       \resizebox{\textwidth}{!}{
-        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
+        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
+        \input{images/frechet_distance/free_space_polygonal_curves.tex}
       }
@@
   % TODO: Replace graphic with appropiate example
-  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \includegraphics[width=0.3\textwidth]{images/frechet_distance/square-ellipse-intersection.pdf}
   \caption[Legal intervals on free space cell boundary]{
-    Legal intervals on the boundaries of a free space cell. 
+    Legal intervals on the boundaries of a free space cell are shown with black white dashed lines. 
     There are at most 8 intersection points for an ellipse and a rectangle.}

### 2022-12-03 docs/paper.tex
refactor

@@
 %%% date
-\date{November 27, 2022}
+\date{December 05, 2022}
 
@@
 
-\section*{Abstract}
+% \section*{Abstract}
 
@@
 It is used in Geo-Information-Systems (GIS) to 
-quantify the similarity or quality of two maps, % cite
+quantify the similarity or quality of two maps \cite{lyu_partial-frechet-distance-based_2022},
 find patterns in vehicle traffic or animal tracking data 
-by trajectory analysis, % cite
-gesture recognition on touch devices like smartphones \cite{hu_research_2022}, % cite
+by trajectory analysis \cite{buchin_detecting_2011},
+gesture recognition on touch devices like smartphones \cite{hu_research_2022},
 and shape matching in the field of computer vision.
@@
 but instead compare curves to curves. 
+Therefore, we start by defining curves in \autoref{def_curve}.
 
@@
 In reality, the course of the curves is important for many applications. 
-For example, in handwriting character recognition on digital input devices, especially in Chinese, % TODO: add more citations
+For example, in handwriting character recognition on digital input devices, especially in Chinese \cite{chanin_david_hanzi_nodate},
 gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
@@
 The length of the dog leash is then exactly the Fréchet distance of the two curves.
-A formal definition is given in \autoref{def_frechet_distance}.
+A formal definition is given in \autoref{def_frechet_distance} 
+along with a visual example in \autoref{fig_frechet_distance_example}.
 
@@
 
+\begin{figure}[ht]
+  \centering
   % TODO: Add example graphic of Frechet distance
+  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \caption{The Fréchet distance is indicated by the black lines connecting the two polygonal curves.}
+  \label{fig_frechet_distance_example}
+\end{figure}
 
+% TODO: merge with above mentions of applications?
 Besides the previously mentioned applications, 
@@
 
-\subsection{Fréchet distance with speed limits}
+% \subsection{Fréchet distance with speed limits}
 
-\subsection{Fréchet Distance for Uncertain Curves}
+% \subsection{Fréchet Distance for Uncertain Curves}
 
@@
   \centering
-  % TODO: update figure graphics
   \begin{subfigure}[b]{0.45\textwidth}
@@
   \centering
-  % TODO: Replace graphic with appropiate example
   \includegraphics[width=0.3\textwidth]{images/frechet_distance/square-ellipse-intersection.pdf}
@@
   \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
-  \caption{The monotonically increasing curve through the free space is computed by dynamic programming.}
+  \caption{The monotonically increasing curve (red) through the free space diagram.}
   \label{fig_free_space_curve}
@@
 
-\subsection{Improved Version?}
-\label{sec_improved_algorithm}
-\cite{van_leusden_novel_2013}
-\cite{buchin_four_2017}
+% \subsection{Improved Version?}
+% \label{sec_improved_algorithm}
+% \cite{van_leusden_novel_2013}
+% \cite{buchin_four_2017}
 
@@
 % Which version to use when
+This work gives an overview on the Fréchet distance, 
+its variants and algorithms to compute them. 
+Several differences between the variants have been described 
+and evaluated. 
+In conclusion, it can be stated that computing the continuous 
+Fréchet distance is more computationally expensive compared 
+to its discrete and weak variants.
+However, computing the continuous variant can be done efficiently
+in practice by binary searching the bits of the desired accuracy directly
+instead of computing the formally correct value. 
+Alternatively, the discrete Fréchet distance can be calculated, 
+since it constrains the continuous distance 
+from below and above. 
+In this way, the tolerance range can be narrowed down.
 

### 2022-12-05 docs/paper.tex
fix typos

@@
   \label{def_hausdorff_distance}
-  Given two (polygonal) curves $P$ and $Q$, 
+  Given two normal or polygonal curves $P$ and $Q$, 
   their Hausdorff distance $\delta_{hd}(P,Q)$ is mathematically defined as: 
@@
 This chapter is structured as follows: 
-first the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
-then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
+first the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm}.
+% then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
 Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_weak_algorithm}.
 Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
-% Finally
 
@@
 We start by looking at the simplest case for two polygonal curves with two vertices each, i.e., two line segments.
-The free space is are all valid combinations of positions on $P$ and $Q$ that have a distance of at most $\varepsilon$.
+The free space is all valid combinations of positions on $P$ and $Q$ that have a distance of at most $\varepsilon$.
 Therefore, we define $F_\varepsilon = \{(s,t) \in [0,1]^2 \ | \ dist(P(s), Q(t)) \leq \varepsilon \}$.

### 2022-12-05 docs/paper.tex
free space monotone curve
add Fréchet distance example figure

@@
 %%%
+
+\usepackage{svg}
 
@@
   % TODO: Add example graphic of Frechet distance
-  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  % \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
+  \includesvg{images/frechet_distance/frechet-distance-example.svg}
   \caption{The Fréchet distance is indicated by the black lines connecting the two polygonal curves.}
@@
   \centering
-  % TODO: Replace graphic with appropiate example
-  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
-  \caption{The monotonically increasing curve (red) through the free space diagram.}
+  \resizebox{0.5\textwidth}{!}{
+    \input{images/frechet_distance/free_space_monotone_curve.tex}
+  }
+  \caption{A monotonically increasing curve (red) through the free space diagram.}
   \label{fig_free_space_curve}

### 2022-12-05 docs/paper.tex
removed name for peer review

@@
   pdftitle={Fréchet Distance}, %%
-  pdfauthor={Tom Stein}, %%
+  % pdfauthor={Tom Stein}, %%
   pdfsubject={Seminar Algorithm Engineering 22/23}, %%
@@
 %%% author(s)
-\author{Tom Stein}
+% \author{Tom Stein}
 

### 2022-12-23 docs/paper.tex
figure as pdf file
Revert "removed name for peer review"
This reverts commit 4843b9037e25fbab2e8b778999e10609656f7d27.
minor (typo) fixes from peer review

--- 2022-12-05 docs/paper.tex
@@
 
-\usepackage{svg}
+% \usepackage{svg}
 
@@
 a distance metric can be used. 
-We can't simply use a distance metric like euclidean distance 
-because we don't want to compare points to points 
+We can not simply use a distance metric like euclidean distance 
+because we do not want to compare points to points 
 but instead compare curves to curves. 
@@
   Given two normal or polygonal curves $P$ and $Q$, 
-  their Hausdorff distance $\delta_{hd}(P,Q)$ is mathematically defined as: 
+  their Hausdorff distance $\delta_{hd}(P,Q)$ is defined as: 
   $$\delta_{hd}(P,Q) = \max \left( \sup_{p \in P} \inf_{q \in Q} \text{dist}(p,q), \sup_{q \in Q} \inf_{p \in P} \text{dist}(p,q) \right)$$
@@
 but only the point sets of the individual curves.
-Two curves may have a small Hausdorff-distance but seem very dissimilar 
+Two curves may have a small Hausdorff distance but seem very dissimilar 
 when looking at them, see \autoref{fig_hausdorff_distance_bad_example}.
@@
 In reality, the course of the curves is important for many applications. 
-For example, in handwriting character recognition on digital input devices, especially in Chinese \cite{chanin_david_hanzi_nodate},
+For example, 
+in handwriting character recognition on digital input devices, especially in Chinese \cite{chanin_david_hanzi_nodate},
+(online) handwriting signature verification \cite{zheng_algorithm_2008,fang_research_2018},
 gesture recognition on smartphone touchscreens \cite{hu_research_2022}, % TODO: add more citations
+map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
+moving object analysis, e.g., detecting commuting patterns \cite{buchin_detecting_2011},
 and motion evaluation in real-time motion capturing \cite{qiao_real-time_2017, shehu_curve_2012}. % TODO: add more citations
@@
   Given two (polygonal) curves $P: [a,a'] \mapsto V$, $Q: [b,b'] \mapsto V$ and a distance function $\text{dist}: (V,V) \mapsto \mathbb{R}$, 
-  their Fréchet distance $\delta_{F}(P,Q)$ is mathematically defined as: 
+  their Fréchet distance $\delta_{F}(P,Q)$ is defined as: 
   $$\delta_{F}(P,Q) = \inf_{\substack{\alpha: [0,1] \mapsto [a, a'] \\ \beta: [0,1] \mapsto [b, b']}} \max_{t \in [0,1]} \text{dist}(P(\alpha(t)), Q(\beta(t)))$$
@@
   % \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
-  \includesvg{images/frechet_distance/frechet-distance-example.svg}
+  \includegraphics{images/frechet_distance/frechet-distance-example.pdf}
   \caption{The Fréchet distance is indicated by the black lines connecting the two polygonal curves.}
@@
 \end{figure}
-
-% TODO: merge with above mentions of applications?
-Besides the previously mentioned applications, 
-where the course of the curves is important, 
-the Fréchet distance also finds applications in
-(online) handwriting signature verification \cite{zheng_algorithm_2008,fang_research_2018},
-map-matching tracking data for traffic control, routing and navigation \cite{wenk_addressing_2006,brakatsoulas_map-matching_2005},
-and moving object analysis, e.g., detecting commuting patterns \cite{buchin_detecting_2011}.
 
@@
 The \textit{weak Fréchet distance} (non-monotone Fréchet distance) 
-slightly modifies the continuous version by allowing going backwards. 
+slightly modifies the continuous version by allowing to go backwards. 
 The definition remains the same as for the continuous Fréchet distance in \autoref{def_frechet_distance}
@@
 The weak Fréchet distance is a lower bound on the continuous Fréchet distance.
-However, the Ratio between them may be arbitrarily large, 
+However, the ratio between them may be arbitrarily large, 
 as can be seen by the structure given in \autoref{fig_weak_frechet_distance}. \cite{alt_computing_1995}
@@
   A \textit{coupling} L between two polygonal curves $P$ and $Q$, 
-  with points $(u_1, \dots, u_p)$ and $(v_1, \dots, v_p)$ respectively, 
+  with points $(u_1, \dots, u_p)$ and $(v_1, \dots, v_q)$ respectively, 
   is defined as a sequence
@@
 
-The discrete variation (\autoref{def_discrete_frechet_distance}) gives an upper bound on the continuous Fréchet distance (\autoref{def_frechet_distance}). 
+The discrete variant (\autoref{def_discrete_frechet_distance}) gives an upper bound on the continuous Fréchet distance (\autoref{def_frechet_distance}). 
 % Clearify if this is really an upper bound for polygonal curves with example
@@
     In the case of the discrete Fréchet distance 
-    one frog has to wait at either start or end $Q$
+    one frog has to wait at either start or end of $Q$
     while the other one is at the middle vertex of $P$, 
@@
 
-Additionally, it's also a lower bound on the continuous Fréchet distance 
+Additionally, it is also a lower bound on the continuous Fréchet distance 
 plus the distance of the longest legal jump of any frog. 
 By increasing the number of points of a polygonal curve that 
-over approximates some other polygonal curve
+overapproximates some other polygonal curve
 this property can be used to compute the continuous Fréchet distance \cite{eiter_computing_1994}.
@@
 Several algorithms to compute the Fréchet distance of polygonal curves were proposed, 
-starting from the Original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
+starting from the original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
 The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions, 
@@
 % then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
-Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_weak_algorithm}.
+Afterwards, the modifications necessary to compute the weak variant are described in \autoref{sec_weak_algorithm}.
 Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
@@
 We start by looking at the simplest case for two polygonal curves with two vertices each, i.e., two line segments.
-The free space is all valid combinations of positions on $P$ and $Q$ that have a distance of at most $\varepsilon$.
+The free space is the set of all valid combinations of positions on $P$ and $Q$ that have a distance of at most $\varepsilon$.
 Therefore, we define $F_\varepsilon = \{(s,t) \in [0,1]^2 \ | \ dist(P(s), Q(t)) \leq \varepsilon \}$.
@@
 These points define the legal regions on each boundary side of the cell, see \autoref{fig_legal_boundary_intervals_of_a_cell}. 
-We don't need to know the exact shape of the free space 
+We do not need to know the exact shape of the free space 
 because we travel in a monotonically increasing way through a convex space,
@@
 the free space of a single cell can be computed in $\mathcal{O}(1)$ time.
-Therefore, the complete free space of all cells can be computed in $\mathcal{O}(pq)$
+Therefore, the complete free space of all cells can be computed in $\mathcal{O}(pq)$ time
 because there are $p-1$ (respectively $q-1$) edges in $P$ (respectively $Q$).
@@
         between two neighboring cells opens up.
-  \item values of $\varepsilon$ where a new passage horizontal (respectively vertical) passage 
+  \item values of $\varepsilon$ where a new horizontal (respectively vertical) passage 
         between two non-neighboring cells opens up.
@@
 \label{sec_weak_algorithm}
-The weak Fréchet distance, as described in \autoref{sec_weak_frechet_distance},
+For the weak Fréchet distance, as described in \autoref{sec_weak_frechet_distance},
 the computation becomes less complex compared to the continuous Fréchet distance.
@@
 one for each cell of the free space diagram,
-and edges connecting the vertices with their, up to six, neighboring vertices.
-Additionally, $s$ is added as start vertex and $t$ as end vertex.
+and edges connecting the vertices with their neighboring vertices.
+There may be up to four neighboring vertices, namely, 
+top, bottom, left, and right.
+Additionally, $s$ is added as the start vertex and $t$ as the end vertex.
 This results in an undirected Graph $G=(V,E)$ 
@@
 with a maximum single edge value of $\varepsilon$.
-This can be checked by removing all edged with values greater than $\varepsilon$ from $E$ 
+This can be checked by removing all edges with values greater than $\varepsilon$ from $E$ 
 and then checking if $s$ is still connected to $t$. 
@@
       dist(u_1, v_1), &\text{if } i=1 \text{and } j=1 \\
-      \max \{c(i-1,j), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
-      \max \{c(i,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
+      \max \{c(i-1,1), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
+      \max \{c(1,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
       \max \{c(i-1,j), c(i,j-1), c(i-1,j-1), dist(u_i, v_j)\}, &\text{if } i>1 \text{and } j>1
@@
 Alternatively, the discrete Fréchet distance can be calculated, 
-since it constrains the continuous distance 
+since it bounds the continuous distance 
 from below and above. 
-In this way, the tolerance range can be narrowed down.
+In this way, the tolerance range can be narrowed down to a desired size.
 

### 2022-12-23 docs/paper.tex
add abstract

@@
 
-% \section*{Abstract}
+\section*{Abstract}
+% The Fréchet distance is a metric used to compare curves in a metric space
+% that takes the course of the curves into account.
+% This report will give an overview on the Fréchet distance 
+% covering the general motivation, 
+% the formal definition along with easy to understand explanations, 
+% the weak and discrete variants, and algorithms to compute all of them.
+
+The Fréchet distance is a metric used to compare curves in a metric space. 
+It measures the similarity between two curves based on the overall shape and course of the curves. 
+This report will provide an overview of the Fréchet distance, 
+including its formal definition, weak and discrete variants, and algorithms for efficiently computing the distance. 
+The report will also cover the general motivation for using the Fréchet distance 
+and its applications in fields such as pattern recognition, computer vision, and geographic information systems. 
+The report is structured to provide a comprehensive understanding of the Fréchet distance, 
+including easy-to-understand explanations 
+and a discussion of trade-offs between computational complexity and accuracy in the algorithms for computing the distance.
 
@@
 is a very theoretical problem, but it has many, not that obvious, practical applications.
-It is used in Geo-Information-Systems (GIS) to 
+It is used in geographic information systems (GIS) to 
 quantify the similarity or quality of two maps \cite{lyu_partial-frechet-distance-based_2022},

### 2022-12-25 docs/paper.tex
add citations to figures

@@
     for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
-    and distance $\varepsilon = 0.73$ using euclidean distance.}
+    and distance $\varepsilon = 0.73$ using euclidean distance. (Figure inspired by \cite{buchin_four_2017})}
   \label{fig_complex_free_space}
@@
     Legal intervals on the boundaries of a free space cell are shown with black white dashed lines. 
-    There are at most 8 intersection points for an ellipse and a rectangle.}
+    There are at most 8 intersection points for an ellipse and a rectangle. (Figure inspired by \cite{alt_computing_1995})}
   \label{fig_legal_boundary_intervals_of_a_cell}
@@
   }
-  \caption{A monotonically increasing curve (red) through the free space diagram.}
+  \caption{A monotonically increasing curve (red) through the free space diagram. 
+    (Figure inspired by \cite{buchin_four_2017})}
   \label{fig_free_space_curve}

### 2022-12-26 docs/paper.tex
display correct diagram grid in free space plots

@@
       \resizebox{\textwidth}{!}{
-        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
-        \input{images/frechet_distance/free_space_polygonal_curves.tex}
+        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
+        % \input{images/frechet_distance/free_space_polygonal_curves.tex}
       }
@@
   \resizebox{0.5\textwidth}{!}{
-    \input{images/frechet_distance/free_space_monotone_curve.tex}
+    \includegraphics[width=\textwidth]{images/frechet_distance/free_space_monotone_curve.pdf}
+    % \input{images/frechet_distance/free_space_monotone_curve.tex}
   }

### 2022-12-26 docs/paper.tex
docs: fixed some errors and updated date.

@@
 %%% date
-\date{December 05, 2022}
+\date{December 26, 2022}
 
@@
       }
-      \caption{Line segments $P$ and $Q$}
+      \caption{Line segments $P$ and $Q$.}
       \label{fig_free_space_lines_example}
@@
       }
-      \caption{Free space diagram}
+      \caption{Free space diagram.}
       \label{fig_free_space_diagram_example}
@@
     The free space diagram \autoref{fig_free_space_diagram_example} 
-    for line segments $P$ and $Q$ and distance $\varepsilon$
+    for line segments $P$ and $Q$ and distance $\varepsilon = 0.73$
     from \autoref{fig_free_space_lines_example} using euclidean distance.}
@@
       }
-      \caption{Polygonal curves $P$ and $Q$}
+      \caption{Polygonal curves $P$ and $Q$.}
       \label{fig_complex_free_space_lines_example}
@@
       }
-      \caption{Free space diagram}
+      \caption{Free space diagram for a non-minimal value of $\varepsilon$.}
       \label{fig_complex_free_space_diagram_example}
@@
 and stopping as soon as $t$ is included in the spanning tree. 
-The value of $\varepsilon$ is the value of the last added edge.
+The value of $\varepsilon$ is the maximum value of all edged added to the spanning tree.
 Similarly, one can use Dijkstra's algorithm \cite{dijkstra_note_1959}, 
@@
 In this case of a planar grid based graph with non-negative edge values 
-an even faster algorithm from Henzinger et al. \cite{henzinger_faster_1997} to solve the single source shortest paths 
+an even faster algorithm from Henzinger et al. \cite{henzinger_faster_1997} to solve the single source shortest paths problem 
 can be applied. This gives a time complexity of $\mathcal{O}(pq)$ to compute the weak Fréchet distance \cite{van_leusden_novel_2013}.
@@
 all other values of $c(i,j)$ are computed iteratively in a bottom up manner 
-and memoization on the previously computed values of $c(i,j)$.
+with memoization on the previously computed values of $c(i,j)$.
 The resulting algorithm has runtime of $\mathcal{O}(pq)$
 because the constant time function $c(i,j)$ is called $pq$ times \cite{eiter_computing_1994}.
+Additionally, the algorithm can be implemented to require $\mathcal{O}(min(p,q))$ space 
+because it is sufficient to store the previous column or row of $c(i,j)$ values 
+for memoization.
 
@@
   \label{def_recursive_forumla_discrete_frechet_distance}
-  The discrete Fréchet distance of polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
+  The discrete Fréchet distance of the two (partial) polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
   can be computed by
